# Results

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(lubridate)
  library(ggplot2)
  library(fs)
  library(ggridges)
  library(dplyr)
  library(tsibble)
  library(reshape2)
  library(fable)
  library(feasts)
  library(urca)
})

source("scripts/DailySummary.R")
source("scripts/WeeklyMonthlyTrend.R")
source("scripts/TimeSeriesDecomp.R")
source("scripts/EventChains.R")
source("scripts/RemoveOutliers.R")
source("scripts/SAIDI.R")
source("scripts/SAIFI.R")
source("scripts/CAIDI.R")


data_path   <- fs::path("data", "PackageData.csv")
trouble_codes_path <- fs::path("data", "TroubleCodeLegend.csv")

outages_raw <- readr::read_csv(data_path, show_col_types = FALSE)
trouble_codes <- readr::read_csv(trouble_codes_path, show_col_types = FALSE)

daily_summary <- summarize_daily_outages(outages_raw)

outages <- prepare_outages(outages_raw)

outages <- outages |> 
  mutate(month = month(outages$STARTDATE))

monthly_summary <- summarise_monthly_outages(outages)

daily_ts <- prepare_daily_ts(outages_raw)

chains_df <- identify_event_chains(outages_raw,
                                   min_gap = 5,
                                   max_gap = 10)
chain_summary <- summarise_event_chains(chains_df)

```

## Daily Outage Summary
### Total outages per day
```{r}
plot_daily_outages(daily_summary,
                     start_date = as.Date("2025-06-01"),  
                     end_date   = as.Date("2025-07-01"))

```

### Total customers affected per day
```{r}
plot_daily_customers(daily_summary,
                     start_date = as.Date("2024-12-01"),  
                     end_date   = as.Date("2025-01-01"))

```

### Total duration of outages per day
```{r}
plot_daily_total_duration(daily_summary,
                     start_date = as.Date("2025-04-01"),  
                     end_date   = as.Date("2025-05-01"))

```

### Average time to restoration
```{r}
plot_daily_avg_restoration(daily_summary,
                     start_date = as.Date("2025-07-01"),  
                     end_date   = as.Date("2025-08-01"))
```

### Percent of momentary events
```{r}
plot_daily_pct_momentary(daily_summary,
                     start_date = as.Date("2025-09-01"),  
                     end_date   = as.Date("2025-10-01"))
```

## Trend Analysis
### Outages per month
```{r}
plot_outages_per_month(monthly_summary)
```

### Customer-hours lost
```{r}
plot_customer_hours_monthly(monthly_summary)
```

### Troubleshooting code patterns
```{r}
plot_troublecode_top_monthly(outages, top_n_codes = 5)
```

### Seasonal outage frequency
```{r}
plot_seasonal_outage_frequency(outages)
```

### Yearly Trend
```{r}
plot_outage_trend_loess(daily_ts)
```

## Event Chains
### Chain counts per borough
```{r}
plot_chain_counts_by_boro(chain_summary)
```

### Share of events in chains (%)
```{r}
plot_chain_percent_by_boro(chain_summary)
```

### Time-of-day pattern (When do chains happen?)
```{r}
plot_chain_time_of_day(chains_df)
```

## Utility Metrics
### SAIDI Counts per Month
#### System Average Interruption Duration Index
```{r}

outages <- outages |>
  filter(XBORO != '-NDA-') |>
  filter(OUTAGEFLAG == 1) |>
  group_by(XBORO) |>
  filter(remove_outliers(TOTALCUSTAFFECTED)) |>
  ungroup()

saidi_list <- split(outages, outages$month)

SAIDI_calc_per_month <- lapply(saidi_list, function(x){
  SAIDI_calc(x)
})

plot_saidi <- data.frame(
  month = names(SAIDI_calc_per_month),
  saidi = unlist(SAIDI_calc_per_month)
)
x_ordered <- factor(plot_saidi$month, 
                    levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))

ggplot(plot_saidi, aes(x=x_ordered, y= saidi)) +
  geom_point() +
  labs(x = "Month", y = "SAIDI",
       title = "SAIDI Calculation per Month")

SAIDI_calc(outages)
```

### SAIFI
#### System Average Interruption Frequency Index
```{r}

SAIFI_calc(outages)

```

### CAIDI
#### Customer Average Interruption Duration Index 

```{r}

CAIDI_calc(outages)

```

## Borough-Level Analysis

### Outage Counts per Borough
```{r}

ggplot(outages, aes(x=XBORO))+
  geom_bar()+
  labs(x="Borough", y="Count of Outages", title ="Outage Count from Oct 2024 to Oct 2025 by Borough")

```

### Boxplot Breakdown per Borough

```{r}
outages_boro <- outages |>
  mutate(
    start_ts = lubridate::ymd_hms(STARTDATE, tz = "UTC"),
    rest_ts  = lubridate::ymd_hms(RESTDATE,  tz = "UTC"),
    
    # Use start time as the "event day" for aggregation
    outage_day = as.Date(start_ts),
    duration_mins = as.numeric(difftime(rest_ts, start_ts, units = "mins"))
  )

quartiles <- outages_boro |>
  group_by(XBORO) |>
  summarise(
    Q1 = quantile(duration_mins, 0.25),
    Median = median(duration_mins),
    Q3 = quantile(duration_mins, 0.75),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(cols = c(Q1, Median, Q3),
                      names_to = "stat",
                      values_to = "y")

ggplot(outages_boro, aes(x = XBORO, y = duration_mins)) +
  geom_boxplot(outliers = FALSE) +
  labs(title="Outage Duration by Borough",
       x= "Borough",
       y="Duration in Minutes") +
  geom_text(
    data = quartiles,
    aes(x = XBORO, y = y, label = round(y, 2)),
    vjust = -0.5,  # Adjust vertical position
    color = "black",
    size = 2
  ) +
  theme(
    axis.text.x = element_text(size = 8)
  )
```

### Ridgeline Plot of Outage Counts per Borough
```{r}

ggplot(outages, aes(x = TOTALCUSTAFFECTED, y = XBORO)) +
geom_density_ridges(scale = 1.2, rel_min_height = 0.01, fill = "lightblue", alpha = 0.5) +
theme_ridges() +
theme(legend.position = "none") +
labs(title = "Ridgeline Plots of Total Cust Affected by Borough",
     x = "Total Customers Affected", 
     y = "Borough")
```

## Trouble Code Analysis

### Facet Bar Graphs for Top 5 Boroughs
```{r}
events_merged <- outages |>
  left_join(trouble_codes, by = c("TROUBLECODE" = "SHORT_DESC"))

top_counts <- events_merged |>
  filter(XBORO != '-NDA-') |>
  filter(CODE_NAME != 'NA') |>
  group_by(XBORO) |>
  count(CODE_NAME, name = "count") |>  
  arrange(XBORO, desc(count), CODE_NAME) |>
  slice_head(n = 5)


ggplot(top_counts, aes(x=CODE_NAME, y=count)) +
    geom_bar(stat="identity") +
    facet_wrap(~XBORO) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(title="Outage Count by Trouble Code",
         subtitle = "References Top 5 Trouble Codes in each Borough",
         x="Trouble Code",
         y="Outage Count")

```


### Cleveland Dot Plot

```{r}
ggplot(top_counts, aes(x=count, y=CODE_NAME)) +
    geom_point() +
    facet_wrap(~XBORO) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(title="Outage Count by Trouble Code",
         subtitle = "References Top 5 Trouble Codes in each Borough",
         x="Trouble Code",
         y="Outage Count")
```
### Heat Map for Top 5 Trouble Codes per Borough
```{r}
ggplot(top_counts, aes(x = CODE_NAME, y = XBORO)) +
    geom_tile(aes(fill = count)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    scale_fill_continuous(high = "#132B43", low = "#56B1F7") +
    labs(title = "Top Troublecode by Boroughs", 
         x = "Trouble Code", 
         y = "Borough")

```
